\chapter{Acknowledgements}

首先，我想感謝在中研院的程式語言與形式方法實驗室服務的陳郁方教授與王柏堯教授。這篇碩士論文是在他們兩位的指導下完成的，在攻讀碩士學位的期間，兩位教授讓我了解到什麼才是「做學問」的態度，也讓我認識了更多在之前從未想像過的問題與方法。另外，也很感謝陳郁方教授願意帶我一起去2016年的ICSE，見證一下頂級會議的規模跟樣貌。除此之外，我也想感謝目前人在UIUC攻讀博士學位的謝橋學長，如果沒有您的話，我的碩士之路應該會走得非常不同。我想，對軟體工程的興趣應該是在您帶領我進入中研院後，才真的被培養出來的。也很感謝同樣在中研院資訊所的蔡明憲博士，您堅強的程式撰寫能力跟對事情的分析能力一直是我難以望其項背的。如果沒有您的協助，我想這篇論文的實作部分應該會薄弱很多。

在台灣大學的軟體測試實驗室之中，我也得到了非常多的收穫。首先要感謝實驗室的大家長--王凡教授，您對於實驗室的同學們總是以耐心相待，而不會過於苛責，大家一齊專心致志地朝著目標前進。也很感謝實驗室的修博、啟允、上誼跟文起，在實驗室的時候可以跟你們聊天、吃零食跟（偶而）摸魚，都是很快樂的回憶。實驗室的助理志奇也替我們處理了不少瑣事，謝謝你總是會提前考慮到許多我們沒有想到的事情，是我們跟學校諸多雜事之間的一道替我們把關的保險。

接著，我想謝謝我的家人。謝謝爸、媽總是關心我有沒有吃飽、睡好，抱歉這兩年沒有很常回家跟大家吃飯、聊自己的近況。另外，爸總提醒我要多運動，可惜我碩士兩年以來，除了對軟體工程多了不少心得之外，體重也是直線上升。農藝系的弟弟常拿巨大的蔬菜回家給大家吃，希望你在往後的大學生活中能夠像現在一樣享受自己所學。最後，我想感謝我的女友，目前剛升上碩士班二年級的趙正媛。有你的陪伴，這兩年才能如此平順的面對生活中各種出乎意料的事情。感謝你總是願意忍受我跳躍的思考與起伏不定的情緒，也謝謝你願意聽完我扯了一堆關於研究、實驗的事情後，拍拍我對我說「雖然我聽不懂，不過沒問題的」。希望你在攻讀碩士的過程中可以找到平穩的步調，也希望我可以當你的依靠。

\vspace{8mm}
\noindent
感謝你們！

\vspace{8mm}

\noindent
National Taiwan University, Taipei

\noindent
July, 2016 \hfill 李宗儒 \ Lii, Tsung-Ju

\begin{comment}
I should like to thank Prof. Yu-Fang Chen and Prof. Bow-Yaw Wang at Formal
Methods Laboratory in Academia Sinica.
Your guidance and instructions helped me gradually develop the whole framework
since I was a full-time research assistant in FM Lab.
Your solid training and teaching also forced me to establish my knowledge of
Software Verification and Formal Methods.
In particular, the opportunity and the experience you gave me to attend
international conferences surely is very enlightening though quite frightening.
In addition, I am specially grateful to and surprised by the assistance from
Dr. Ming-Hsien Tsai on the implementation.
Your efficiency and proficiency on software programming skills are simply
remarkable and incomparable.
Without your assistance, I am sure I cannot finish the implementation on my own,
let alone participating the competition.

I also learned a lot and pretty enjoyed my life in Software Testing Laboratory
in National Taiwan University.
Prof. Farn Wang is a great role model as a academic researcher.
His ambition on building an industrial strength software testing tool is
respectable.
The laboratory members and assistants are just adorable.
You guys inspired me in a unique way.
Even when we were both struggling for homeworks, reports, exams, and finally
the graduation, you guys still found many ways to relax and relief the tension
among the whole lab.
"Work Hard, Play Hard" is the spirit I learned from you.
You guys rock.

Finally, my appreciation to my parents and siblings is beyond words.
Thanks for your constant nagging on me about my physically and mentally
unhealthy lifestyle as well as trying to stop my 26-year-without-girlfriend
achievement.
Your concerns and considerations kept reminding me that,
even if I might fail so hard and be so frustrated, I still can go home,
where I can relax and recover until the next challenge comes.

\end{comment}


